---
title: "Lineup Projector + Optimizer"
author: "Dylan Kinion"
date: "2024-07-06"
output: 
  html_document:
    code_folding: hide
    toc: true
runtime: shiny
---
Possible ways to improve:
* features
  * add height/measurables to oreb_pct
  * ftr and to rate are over reliant on own stats (i.e ftr relies to heavily on individual ftr)
* aggregation method for models


Notes on model:

packages 
```{r, error=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(lpSolve)
library(jsonlite)
library(lubridate)
library(rsample)
library(parsnip)
library(caret)
library(class)
library(Metrics)
library(shiny)
library(rsconnect)
library(rvest)
library(stats)
library(plotly)
library(reshape2)
library(DT)
library(scales)

Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)
options(rsconnect.max.bundle.size = 5000*1024^2)
```

player stats
```{r, error=FALSE, warning=FALSE, message=FALSE}
get_season_data <- function(season) {
  base_url <- "https://www.basketball-reference.com/leagues/NBA_"
  
  # URLs for the per-minute and advanced stats pages
  url_per_minute <- paste0(base_url, season, "_per_minute.html")
  url_advanced <- paste0(base_url, season, "_advanced.html")
  
  # Scrape per-minute stats
  stats <- read_html(url_per_minute) %>%
    html_node("table") %>%
    html_table(fill = TRUE) %>%
    filter(Player != "Player") %>%
    select(-Rk)
  
  # Scrape advanced stats
  adv_stats <- read_html(url_advanced) %>%
    html_node("table") %>%
    html_table(fill = TRUE) %>%
    select(2:19, 21:24, 26:29) %>%
    filter(Player != "Player")
  
  # Combine the stats based on Player, Position, and Team
  combined_stats <- right_join(stats, adv_stats, by = c("Player", "Pos", "Tm"))
  
  # Add a column identifying the season
  combined_stats <- combined_stats %>%
    mutate(Season = season)
  
  return(combined_stats)
}

# Loop through the seasons and get data
get_multiple_seasons_data <- function(start_year, end_year) {
  season_data <- list()
  
  for (season in start_year:end_year) {
    print(paste("Getting data for season:", season))
    season_data[[as.character(season)]] <- get_season_data(season)
  }
  
  return(season_data)
}

# Get data for seasons from 2015 to 2024
all_seasons_data <- get_multiple_seasons_data(2015, 2024)
combined_df <- bind_rows(all_seasons_data)
library(stringi)
combined_df$Player <- stri_trans_general(combined_df$Player, "Latin-ASCII")

combined_df$`3P%` <- as.numeric(combined_df$`3P%`)
combined_df$`FT%` <- as.numeric(combined_df$`FT%`)
combined_df$`TS%` <- as.numeric(combined_df$`TS%`)
combined_df$`3PAr` <- as.numeric(combined_df$`3PAr`)
combined_df$`FTr` <- as.numeric(combined_df$`FTr`)
combined_df$`TRB%` <- as.numeric(combined_df$`TRB%`)
combined_df$`BLK%` <- as.numeric(combined_df$`BLK%`)
combined_df$`AST%` <- as.numeric(combined_df$`AST%`)
combined_df$`TOV%` <- as.numeric(combined_df$`TOV%`)
combined_df$`ORB%` <- as.numeric(combined_df$`ORB%`)
combined_df$`DRB%` <- as.numeric(combined_df$`DRB%`)
combined_df$`USG%` <- as.numeric(combined_df$`USG%`)
combined_df$`FGA` <- as.numeric(combined_df$`FGA`)
combined_df$`PTS` <- as.numeric(combined_df$`PTS`)
combined_df$`MP.x` <- as.numeric(combined_df$`MP.x`)
combined_df$`AST` <- as.numeric(combined_df$`AST`)
combined_df$`TOV` <- as.numeric(combined_df$`TOV`)

player_pos <- combined_df %>%
  mutate(groupPosition = case_when(
    Pos == "PF" ~ 2,
    Pos == "SG" ~ 2,
    Pos == "C" ~ 3,
    Pos == "SF" ~ 2,
    Pos == "PG" ~ 1,
    Pos == "SG-PG" ~ 1,
    Pos == "PG-SG" ~ 1,
    Pos == "SF-PF" ~ 2,
    Pos == "SF-SG" ~ 2,
    Pos == "SG-SF" ~ 2,
    Pos == "PF-SF" ~ 2,
    Pos == "PF-C" ~ 1,
    Pos == "SG-PF" ~ 2,
    Pos == "C-PF" ~ 1,
    Pos == "SF-C" ~ 2,
    Pos == "SG-PG-SF" ~ 1,
    TRUE ~ 2
  )) %>%
  select(Player, Tm, Season, groupPosition)

player_df <- combined_df %>%
  filter(MP.x > 50) %>%
  mutate(`TRB%` = `TRB%` * .01,
         `AST%` = `AST%` * .01,
         `BLK%` = `BLK%` * .01,
         `USG%` = `USG%` * .01,
         `TOV%` = `TOV%` * .01,
         `ORB%` = `ORB%` * .01) %>%
  mutate(pps = PTS/FGA,
         ast_to = AST/TOV) %>%
  group_by(Player, Season) %>%
  mutate(avg_pct_fg3 = (`3P%` + ifelse(is.na(lag(`3P%`)), 0.32, lag(`3P%`))) / 2,
         avg_3pt_rate = (`3PAr` + ifelse(is.na(lag(`3PAr`)), 0.36, lag(`3PAr`))) / 2,
         avg_TS = (`TS%` + ifelse(is.na(lag(`TS%`)), 0.53, lag(`TS%`))) / 2,
         avg_ft_pct = (`FT%` + ifelse(is.na(lag(`FT%`)), 0.74, lag(`FT%`))) / 2,
         avg_ft_rate = (`FTr` + ifelse(is.na(lag(`FTr`)), 0.26, lag(`FTr`))) / 2,
         avg_treb = (`TRB%` + ifelse(is.na(lag(`TRB%`)), 0.09, lag(`TRB%`))) / 2,
         avg_ast_pct = (`AST%` + ifelse(is.na(lag(`AST%`)), 0.13, lag(`AST%`))) / 2,
         avg_blk_pct = (`BLK%` + ifelse(is.na(lag(`BLK%`)), 0.02, lag(`BLK%`))) / 2,
         avg_usage = (`USG%` + ifelse(is.na(lag(`USG%`)), 0.19, lag(`USG%`))) / 2,
         avg_tov_pct = (`TOV%` + ifelse(is.na(lag(`TOV%`)), 0.012, lag(`TOV%`))) / 2,
         avg_oreb_pct = (`ORB%` + ifelse(is.na(lag(`ORB%`)), 0.05, lag(`ORB%`))) / 2,
         avg_dreb_pct = (`DRB%` + ifelse(is.na(lag(`DRB%`)), 0.149, lag(`DRB%`))) / 2,
         avg_pps = (pps + ifelse(is.na(lag(pps)), 1.208, lag(pps))) / 2,
         avg_ast_to = (ast_to + ifelse(is.na(lag(ast_to)), 1.778, lag(ast_to))) / 2
           ) %>%
  ungroup() %>%
  na.omit()

t <- player_df %>%
  filter(!is.na(ast_to),
         ast_to != "Inf",
         MP.x > 100) 
mean(t$ast_to)

player_df$avg_pps <- rescale(player_df$avg_pps)
player_df$avg_ast_to <- rescale(player_df$avg_ast_to)

select_stats <- right_join(player_df, player_pos, by = c("Player", "Season")) %>%
  rename(namePlayer = Player,
         slugTeamBREF = Tm.x) %>%
  mutate(pred_season = Season + 1) %>%
  select(namePlayer,
         slugTeamBREF,
         avg_pct_fg3,
         avg_3pt_rate,
         avg_TS,
         avg_ft_pct,
         avg_ft_rate,
         avg_treb,
         avg_ast_pct,
         avg_blk_pct,
         avg_tov_pct,
         avg_usage,
         avg_oreb_pct,
         avg_dreb_pct,
         avg_pps,
         avg_ast_to,
         groupPosition,
         pred_season)
stats <- right_join(player_df, player_pos, by = c("Player", "Season")) %>%
  rename(namePlayer = Player,
         teamAbbreviation = Tm.x,
         yearSeason = Season) %>%
  select(namePlayer,
         teamAbbreviation,
         avg_pct_fg3,
         avg_3pt_rate,
         avg_TS,
         avg_ft_pct,
         avg_ft_rate,
         avg_treb,
         avg_ast_pct,
         avg_blk_pct,
         groupPosition,
         yearSeason,
         avg_usage,
         avg_pps,
         avg_ast_to)

```

```{r, error=FALSE, warning=FALSE, message=FALSE}

```

lineup stats
```{r, error=FALSE, warning=FALSE, message=FALSE}
seasons <- c("2023-24", "2022-23", "2021-22", "2020-21", "2019-20", "2018-19", "2017-18", "2016-17", "2015-16", "2014-15")

# Initialize an empty list to store data frames
all_seasons_data <- list()

# Loop through each season and fetch the data
for (season in seasons) {
  url <- paste0("https://api.pbpstats.com/get-totals/nba?Season=", season, "&SeasonType=Regular%2BSeason&Type=Lineup")
  data <- fromJSON(url)
  lineup_data <- data[["multi_row_table_data"]]
  
  # Convert to data frame, add season column
  lineup_data <- as.data.frame(lineup_data)
  lineup_data$season <- season  # Add season year
  
  # Store the data frame in the list
  all_seasons_data[[season]] <- lineup_data
}

# Combine all seasons data into one data frame
combined_data <- bind_rows(all_seasons_data)
combined_data <- combined_data %>%
  filter(!is.na(PlusMinus))

opp_seasons_data <- list()

# Loop through each season and fetch the data
for (season in seasons) {
  url_opp <- paste0("https://api.pbpstats.com/get-totals/nba?Season=", season, "&SeasonType=Regular%2BSeason&Type=LineupOpponent")
  data_opp <- fromJSON(url_opp)
  lineup_data_opp <- data_opp[["multi_row_table_data"]]
  
  # Convert to data frame, add season column
  lineup_data_opp <- as.data.frame(lineup_data_opp)
  lineup_data_opp$season <- season  # Add season year
  
  # Store the data frame in the list
  opp_seasons_data[[season]] <- lineup_data_opp
}

# Combine all seasons data into one data frame
combined_data_opp <- bind_rows(opp_seasons_data)
combined_data_opp <- combined_data_opp %>%
  filter(!is.na(PlusMinus)) %>%
  rename(opp_def_rebs = DefRebounds) %>%
  select(EntityId,
         opp_def_rebs)
combined_data2 <- right_join(combined_data, combined_data_opp, by = "EntityId")

lineup_data <- combined_data2 %>%
  filter(OffPoss > 150) %>%
  separate(Name, into = c("Player 1", "Player 2", "Player 3", "Player 4", "Player 5"), sep = ", ") %>%
  mutate( fga = FG2A + FG3A,
    to_pct = Turnovers/(fga + (.44 * FTA) + Turnovers),
    off_reb_pct = OffRebounds/ (OffRebounds + opp_def_rebs),
    ftr = FTA/fga) %>%
  select(TeamAbbreviation,
         season,
         EfgPct,
         to_pct,
         off_reb_pct,
         ftr,
         `Player 1`,
         `Player 2`,
         `Player 3`,
         `Player 4`,
         `Player 5`) %>%
  mutate(pred_season = case_when(
    season == "2023-24" ~ "2024",
    season == "2022-23" ~ "2023",
    season == "2021-22" ~ "2022",
    season == "2020-21" ~ "2021",
    season == "2019-20" ~ "2020",
    season == "2018-19" ~ "2019",
    season == "2017-18" ~ "2018",
    season == "2016-17" ~ "2017",
    season == "2015-16" ~ "2016",
    TRUE ~ "2015"
  )) %>%
  select(-c(season))
lineup_data$pred_season <- as.numeric(lineup_data$pred_season)
```

combining data frames
```{r, error=FALSE, warning=FALSE, message=FALSE}
df_if <- lineup_data %>% 
  pivot_longer(
    cols = starts_with("Player"),   # Select the player columns
    names_to = "Player",            # Create a new column for the player identifier
    values_to = "namePlayer"        # Create a new column for player names
  ) %>%
  mutate(yearSeason = pred_season - 1)

df_stats <- right_join(df_if, select_stats, by = c("namePlayer", "pred_season")) %>%
  na.omit() %>%
  distinct() 

df_efg <- df_stats %>%
  group_by(TeamAbbreviation, EfgPct, pred_season) %>%
  summarize(
    avg_3pt_rate = mean(avg_3pt_rate),
    min_3pt_rate = min(avg_3pt_rate),
    max_3pt_rate = max(avg_3pt_rate),
    avg_pct_fg3 = mean(avg_pct_fg3),
    min_pct_fg3 = min(avg_pct_fg3),
    max_pct_fg3 = max(avg_pct_fg3),
    avg_ft_pct = mean(avg_ft_pct),
    min_ft_pct = min(avg_ft_pct),
    max_ft_pct = max(avg_ft_pct),
    avg_TS = mean(avg_TS),
    min_TS = min(avg_TS),
    max_TS = max(avg_TS),
    avg_pps = mean(avg_pps),
    min_pps = min(avg_pps),
    max_pps = max(avg_pps)
  ) %>%
  ungroup()

df_to_rate <- df_stats %>%
  filter(avg_ast_to != "Inf") %>%
  group_by(TeamAbbreviation, to_pct, pred_season) %>%
  summarize(
    avg_ast_pct = mean(avg_ast_pct),
    min_ast_pct = min(avg_ast_pct),
    max_ast_pct = max(avg_ast_pct),
    avg_tov_pct = mean(avg_tov_pct),
    min_tov_pct = min(avg_tov_pct),
    max_tov_pct = max(avg_tov_pct),
    avg_usage = mean(avg_usage),
    min_usage = min(avg_usage),
    max_usage = max(avg_usage),
    avg_ast_to = mean(avg_ast_to),
    min_ast_to = min(avg_ast_to),
    max_ast_to = max(avg_ast_to)
  ) %>%
  ungroup()
  
df_oreb <- df_stats %>%
  group_by(TeamAbbreviation, off_reb_pct, pred_season) %>%
  summarize(
    avg_dreb_pct = mean(avg_dreb_pct),
    min_dreb_pct = min(avg_dreb_pct),
    max_dreb_pct = max(avg_dreb_pct),
    avg_oreb_pct = mean(avg_oreb_pct),
    min_oreb_pct = min(avg_oreb_pct),
    max_oreb_pct = max(avg_oreb_pct)
  ) %>%
  ungroup()

df_ftr <- df_stats %>%
  group_by(TeamAbbreviation, ftr, pred_season) %>%
  summarize(
    avg_ftr = mean(avg_ft_pct),
    min_ftr = min(avg_ft_rate),
    max_ftr = max(avg_ft_rate),
    avg_usage = mean(avg_usage),
    min_usage = min(avg_usage),
    max_usage = max(avg_usage),
    avg_ft_pct = mean(avg_ft_pct),
    min_ft_pct = min(avg_ft_pct),
    max_ft_pct = max(avg_ft_pct)
  ) %>%
  ungroup()
```

data splitting
```{r, error=FALSE, warning=FALSE, message=FALSE}
set.seed(3445)

efg_split <- initial_split(df_efg, prop = .90, strata = EfgPct)

efg_train <- efg_split %>%
  training() 
efg_train_c <- efg_train %>%
  select(-c(1,3))

efg_test <- efg_split %>%
  testing() 
efg_test_c <- efg_test %>%
  select(-c(1,3))

sd(efg_train$EfgPct)

ftr_split <- initial_split(df_ftr, prop = .90, strata = ftr)

ftr_train <- ftr_split %>%
  training() 
ftr_train_c <- ftr_train %>%
  select(-c(1,3))

ftr_test <- ftr_split %>%
  testing() 
ftr_test_c <- ftr_test %>%
  select(-c(1,3))

sd(ftr_train$ftr)

oreb_split <- initial_split(df_oreb, prop = .90, strata = off_reb_pct)

oreb_train <- oreb_split %>%
  training() 
oreb_train_c <- oreb_train %>%
  select(-c(1,3))

oreb_test <- oreb_split %>%
  testing() 
oreb_test_c <- oreb_test %>%
  select(-c(1,3))

sd(oreb_train$off_reb_pct)

tor_split <- initial_split(df_to_rate, prop = .90, strata = to_pct)

tor_train <- tor_split %>%
  training() 
tor_train_c <- tor_train %>%
  select(-c(1,3))

tor_test <- tor_split %>%
  testing() 
tor_test_c <- tor_test %>%
  select(-c(1,3))

sd(tor_train$to_pct)
```

SVM
```{r, error=FALSE, warning=FALSE, message=FALSE}
svm_model <- svm_linear() %>%
  set_engine("kernlab") %>%
  set_mode("regression")

svm_fit_efg <- svm_model %>%
  fit(EfgPct ~ ., data = efg_train_c)
svm_preds_efg <- predict(svm_fit_efg, new_data = efg_test_c) %>%
  as.data.frame() %>%
  cbind(efg_test$EfgPct)
MAE(svm_preds_efg$.pred, efg_test$EfgPct, na.rm = FALSE)

svm_fit_ftr <- svm_model %>%
  fit(ftr ~ ., data = ftr_train_c)
svm_preds_ftr <- predict(svm_fit_ftr, new_data = ftr_test_c) %>%
  as.data.frame() %>%
  cbind(ftr_test$ftr)
MAE(svm_preds_ftr$.pred, ftr_test$ftr, na.rm = FALSE)

svm_fit_oreb <- svm_model %>%
  fit(off_reb_pct ~ ., data = oreb_train_c)
svm_preds_oreb <- predict(svm_fit_oreb, new_data = oreb_test_c) %>%
  as.data.frame() %>%
  cbind(oreb_test$off_reb_pct)
MAE(svm_preds_oreb$.pred, oreb_test$off_reb_pct, na.rm = FALSE)

svm_fit_tor <- svm_model %>%
  fit(to_pct ~ ., data = tor_train_c)
svm_preds_tor <- predict(svm_fit_tor, new_data = tor_test_c) %>%
  as.data.frame() %>%
  cbind(tor_test$to_pct)
MAE(svm_preds_tor$.pred, tor_test$to_pct, na.rm = FALSE)
```

rand forest
```{r, error=FALSE, warning=FALSE, message=FALSE}
rf_model <- rand_forest() %>%
  set_engine("ranger") %>%
  set_mode("regression")

rf_fit_efg <- rf_model %>%
  fit(EfgPct ~ ., data = efg_train_c)
rf_preds_efg <- predict(rf_fit_efg, new_data = efg_test_c) %>%
  as.data.frame() %>%
  cbind(efg_test$EfgPct)
MAE(rf_preds_efg$.pred, efg_test$EfgPct, na.rm = FALSE)

rf_fit_ftr <- rf_model %>%
  fit(ftr ~ ., data = ftr_train_c)
rf_preds_ftr <- predict(rf_fit_ftr, new_data = ftr_test_c) %>%
  as.data.frame() %>%
  cbind(ftr_test$ftr)
MAE(rf_preds_ftr$.pred, ftr_test$ftr, na.rm = FALSE)

rf_fit_oreb <- rf_model %>%
  fit(off_reb_pct ~ ., data = oreb_train_c)
rf_preds_oreb <- predict(rf_fit_oreb, new_data = oreb_test_c) %>%
  as.data.frame() %>%
  cbind(oreb_test$off_reb_pct)
MAE(rf_preds_oreb$.pred, oreb_test$off_reb_pct, na.rm = FALSE)

rf_fit_tor <- rf_model %>%
  fit(to_pct ~ ., data = tor_train_c)
rf_preds_tor <- predict(rf_fit_tor, new_data = tor_test_c) %>%
  as.data.frame() %>%
  cbind(tor_test$to_pct)
MAE(rf_preds_tor$.pred, tor_test$to_pct, na.rm = FALSE)
```

gbm
```{r, error=FALSE, warning=FALSE, message=FALSE}
gbm_model <- boost_tree() %>%
  set_engine("xgboost") %>%
  set_mode("regression")

gbm_fit_efg <- gbm_model %>%
  fit(EfgPct ~ ., data = efg_train_c)
gbm_preds_efg <- predict(gbm_fit_efg, new_data = efg_test_c) %>%
  as.data.frame() %>%
  cbind(efg_test$EfgPct)
MAE(gbm_preds_efg$.pred, efg_test$EfgPct, na.rm = FALSE)

gbm_fit_ftr <- gbm_model %>%
  fit(ftr ~ ., data = ftr_train_c)
gbm_preds_ftr <- predict(gbm_fit_ftr, new_data = ftr_test_c) %>%
  as.data.frame() %>%
  cbind(ftr_test$ftr)
MAE(gbm_preds_ftr$.pred, ftr_test$ftr, na.rm = FALSE)

gbm_fit_oreb <- gbm_model %>%
  fit(off_reb_pct ~ ., data = oreb_train_c)
gbm_preds_oreb <- predict(gbm_fit_oreb, new_data = oreb_test_c) %>%
  as.data.frame() %>%
  cbind(oreb_test$off_reb_pct)
MAE(gbm_preds_oreb$.pred, oreb_test$off_reb_pct, na.rm = FALSE)

gbm_fit_tor <- gbm_model %>%
  fit(to_pct ~ ., data = tor_train_c)
gbm_preds_tor <- predict(gbm_fit_tor, new_data = tor_test_c) %>%
  as.data.frame() %>%
  cbind(tor_test$to_pct)
MAE(gbm_preds_tor$.pred, tor_test$to_pct, na.rm = FALSE)
```

data for predictions
# adjust data for 2025 rosters, could not find data source for '25 rosters
```{r, error=FALSE, warning=FALSE, message=FALSE}
stats_24 <- select_stats %>%
  filter(pred_season == 2025)

```

```{r, error=FALSE, warning=FALSE, message=FALSE}

player_pool <- stats_24 %>%
  select(-c(pred_season)) %>%
  rename(team = slugTeamBREF) %>%
  filter(team != "TOT") %>%
  distinct(namePlayer, .keep_all = TRUE) %>%
  na.omit()

```

lineup function
```{r, error=FALSE, warning=FALSE, message=FALSE}
lineup_function <- function(player1, player2, player3, player4, player5) {
  lineup <- player_pool %>%
  filter(namePlayer %in% c(player1, player2, player3, player4, player5))
l_efg <- lineup %>%
  summarize(
    avg_3pt_rate = mean(avg_3pt_rate),
    min_3pt_rate = min(avg_3pt_rate),
    max_3pt_rate = max(avg_3pt_rate),
    avg_pct_fg3 = mean(avg_pct_fg3),
    min_pct_fg3 = min(avg_pct_fg3),
    max_pct_fg3 = max(avg_pct_fg3),
    avg_ft_pct = mean(avg_ft_pct),
    min_ft_pct = min(avg_ft_pct),
    max_ft_pct = max(avg_ft_pct),
    avg_TS = mean(avg_TS),
    min_TS = min(avg_TS),
    max_TS = max(avg_TS),
    avg_ast_pct = mean(avg_ast_pct),
    min_ast_pct = min(avg_ast_pct),
    max_ast_pct = max(avg_ast_pct)
  )
efg_pred <- (predict(svm_fit_efg, new_data = l_efg) + predict(rf_fit_efg, new_data = l_efg) +predict(gbm_fit_efg, new_data = l_efg))/3 %>%
  as.data.frame()
efg_pred <- efg_pred %>%
  rename(`EFG%` = `.pred`)

l_ftr <- lineup %>%
  summarize(
    avg_ftr = mean(avg_ft_rate),
    min_ftr = min(avg_ft_rate),
    max_ftr = max(avg_ft_rate),
    avg_usage = mean(avg_usage),
    min_usage = min(avg_usage),
    max_usage = max(avg_usage)
  )
ftr_pred <- (predict(svm_fit_ftr, new_data = l_ftr) + predict(rf_fit_ftr, new_data = l_ftr) + predict(gbm_fit_ftr, new_data = l_ftr))/3 %>%
  as.data.frame()
ftr_pred <- ftr_pred %>%
  rename(`FT Rate` = .pred)

l_oreb <- lineup %>%
  summarize(
    avg_oreb_pct = mean(avg_oreb_pct),
    min_oreb_pct = min(avg_oreb_pct),
    max_oreb_pct = max(avg_oreb_pct),
    avg_dreb_pct = mean(avg_dreb_pct),
    min_dreb_pct = min(avg_dreb_pct),
    max_dreb_pct = max(avg_dreb_pct)
  )
oreb_pred <- (predict(svm_fit_oreb, new_data = l_oreb) + predict(rf_fit_oreb, new_data = l_oreb) + predict(gbm_fit_oreb, new_data = l_oreb))/3 %>%
  as.data.frame()
oreb_pred <- oreb_pred %>%
  rename(`OREB%` = .pred)

l_tor <- lineup %>%
  summarize(
    avg_ast_pct = mean(avg_ast_pct),
    min_ast_pct = min(avg_ast_pct),
    max_ast_pct = max(avg_ast_pct),
    avg_tov_pct = mean(avg_tov_pct),
    min_tov_pct = min(avg_tov_pct),
    max_tov_pct = max(avg_tov_pct),
    avg_usage = mean(avg_usage),
    min_usage = min(avg_usage),
    max_usage = max(avg_usage)
  )
tor_pred <- (predict(svm_fit_tor, new_data = l_tor) + predict(rf_fit_tor, new_data = l_tor) + predict(gbm_fit_tor, new_data = l_tor))/3 %>%
  as.data.frame()
tor_pred <- tor_pred %>%
  rename(`TO Rate` = .pred)

combined_preds <- cbind(efg_pred, ftr_pred) %>%
  cbind(oreb_pred) %>%
  cbind(tor_pred) %>%
  as.data.frame() %>%
  mutate(Sample = "Projected Linuep")

return(combined_preds)
}
lineup_results <- lineup_function("Jaysom Tatum", "Jaylen Brown", "Kristaps Porzingis", "Derrick White", "Jrue Holiday")

chart_function <- function(pred_lineup) {
  
  league_df <- data.frame(
  `EFG%` = mean(lineup_data$EfgPct, na.rm = TRUE),
  `FT Rate` = mean(lineup_data$ftr, na.rm = TRUE),
  `OREB%` = mean(lineup_data$off_reb_pct, na.rm = TRUE),
  `TO Rate` = mean(lineup_data$to_pct, na.rm = TRUE),
  Sample = "League Avg"
) %>%
  rename(`EFG%` = EFG.,
         `FT Rate` = FT.Rate,
         `OREB%` = OREB.,
         `TO Rate` = TO.Rate)

  chart_df <- rbind(league_df, pred_lineup)
  long_df <- melt(chart_df, id.vars = "Sample", 
                measure.vars = c("EFG%", "OREB%", "FT Rate", "TO Rate"), 
                variable.name = "Metric", 
                value.name = "Value")
  
  plot_ly(long_df, x = ~Metric, y = ~Value, color = ~Sample, colors = c("blue", "red"), 
        type = 'bar', barmode = 'group') %>%
  layout(title = "Comparison of League Avg vs Projected Lineup",
         xaxis = list(title = "Metrics"),
         yaxis = list(title = "Value"),
         legend = list(title = list(text = "Sample")))
}
chart_function(lineup_results)

# distance for similar lineups 
dis_df <- lineup_data %>%
  filter(pred_season > 2020)
league_dis <- dis_df %>%
  select(EfgPct,
         to_pct,
         off_reb_pct,
         ftr) %>%
  rename(`EFG%` = EfgPct,
         `FT Rate` = ftr,
         `OREB%` = off_reb_pct,
         `TO Rate` = to_pct)
euclidean_distances <- apply(league_dis, 1, function(row) {
  sqrt(sum((row - lineup_results[ ,-5])^2))
})
euc_res <- cbind(euclidean_distances, dis_df) %>%
  arrange(euclidean_distances) %>%
  head(5) %>%
  mutate(Year = pred_season - 1) %>%
  select(-pred_season)

```
combined function - model, chart, similar lineups
```{r, error=FALSE, warning=FALSE, message=FALSE}
# add team averages using a reference to a player
combined_function <- function(player1, player2, player3, player4, player5) {
  lineup <- player_pool %>%
  filter(namePlayer %in% c(player1, player2, player3, player4, player5))
l_efg <- lineup %>%
  summarize(
    avg_3pt_rate = mean(avg_3pt_rate),
    min_3pt_rate = min(avg_3pt_rate),
    max_3pt_rate = max(avg_3pt_rate),
    avg_pct_fg3 = mean(avg_pct_fg3),
    min_pct_fg3 = min(avg_pct_fg3),
    max_pct_fg3 = max(avg_pct_fg3),
    avg_ft_pct = mean(avg_ft_pct),
    min_ft_pct = min(avg_ft_pct),
    max_ft_pct = max(avg_ft_pct),
    avg_TS = mean(avg_TS),
    min_TS = min(avg_TS),
    max_TS = max(avg_TS),
    avg_pps = mean(avg_pps),
    min_pps = min(avg_pps),
    max_pps = max(avg_pps)
  )
efg_pred <- (predict(svm_fit_efg, new_data = l_efg) + predict(rf_fit_efg, new_data = l_efg) +predict(gbm_fit_efg, new_data = l_efg))/3 %>%
  as.data.frame()
efg_pred <- efg_pred %>%
  rename(`EFG%` = `.pred`)

l_ftr <- lineup %>%
  summarize(
    avg_ftr = mean(avg_ft_rate),
    min_ftr = min(avg_ft_rate),
    max_ftr = max(avg_ft_rate),
    avg_usage = mean(avg_usage),
    min_usage = min(avg_usage),
    max_usage = max(avg_usage),
    avg_ft_pct = mean(avg_ft_pct),
    min_ft_pct = min(avg_ft_pct),
    max_ft_pct = max(avg_ft_pct)
  )
ftr_pred <- (predict(svm_fit_ftr, new_data = l_ftr) + predict(rf_fit_ftr, new_data = l_ftr) + predict(gbm_fit_ftr, new_data = l_ftr))/3 %>%
  as.data.frame()
ftr_pred <- ftr_pred %>%
  rename(`FT Rate` = .pred)

l_oreb <- lineup %>%
  summarize(
    avg_oreb_pct = mean(avg_oreb_pct),
    min_oreb_pct = min(avg_oreb_pct),
    max_oreb_pct = max(avg_oreb_pct),
    avg_dreb_pct = mean(avg_dreb_pct),
    min_dreb_pct = min(avg_dreb_pct),
    max_dreb_pct = max(avg_dreb_pct)
  )
oreb_pred <- (predict(svm_fit_oreb, new_data = l_oreb) + predict(rf_fit_oreb, new_data = l_oreb) + predict(gbm_fit_oreb, new_data = l_oreb))/3 %>%
  as.data.frame()
oreb_pred <- oreb_pred %>%
  rename(`OREB%` = .pred)

l_tor <- lineup %>%
  summarize(
    avg_ast_pct = mean(avg_ast_pct),
    min_ast_pct = min(avg_ast_pct),
    max_ast_pct = max(avg_ast_pct),
    avg_tov_pct = mean(avg_tov_pct),
    min_tov_pct = min(avg_tov_pct),
    max_tov_pct = max(avg_tov_pct),
    avg_usage = mean(avg_usage),
    min_usage = min(avg_usage),
    max_usage = max(avg_usage),
    avg_ast_to = mean(avg_ast_to),
    min_ast_to = min(avg_ast_to),
    max_ast_to = max(avg_ast_to)
  )
tor_pred <- (predict(svm_fit_tor, new_data = l_tor) + predict(rf_fit_tor, new_data = l_tor) + predict(gbm_fit_tor, new_data = l_tor))/3 %>%
  as.data.frame()
tor_pred <- tor_pred %>%
  rename(`TO Rate` = .pred)

combined_preds <- cbind(efg_pred, ftr_pred) %>%
  cbind(oreb_pred) %>%
  cbind(tor_pred) %>%
  as.data.frame() %>%
  mutate(Sample = "Projected Linuep")

 league_avgs <- lineup_data %>%
   filter(pred_season == 2024)
   
 league_df <- data.frame(
  `EFG%` = mean(league_avgs$EfgPct, na.rm = TRUE),
  `FT Rate` = mean(league_avgs$ftr, na.rm = TRUE),
  `OREB%` = mean(league_avgs$off_reb_pct, na.rm = TRUE),
  `TO Rate` = mean(league_avgs$to_pct, na.rm = TRUE),
  Sample = "League Avg - 2024"
) %>%
  rename(`EFG%` = EFG.,
         `FT Rate` = FT.Rate,
         `OREB%` = OREB.,
         `TO Rate` = TO.Rate)
  
  team_avg <- league_avgs %>%
    filter(TeamAbbreviation == lineup$team[1]) %>%
    summarize(
      `EFG%` = mean(EfgPct, na.rm = TRUE),
      `FT Rate` = mean(ftr, na.rm = TRUE),
      `OREB%` = mean(off_reb_pct, na.rm = TRUE),
      `TO Rate` = mean(to_pct, na.rm = TRUE),
    ) %>%
    mutate(Sample = "Team Avg - 2024")
  chart_df <- rbind(league_df, combined_preds, team_avg)
  long_df <- melt(chart_df, id.vars = "Sample", 
                measure.vars = c("EFG%", "OREB%", "FT Rate", "TO Rate"), 
                variable.name = "Metric", 
                value.name = "Value")
  
  chart <- plot_ly(long_df, x = ~Metric, y = ~Value, color = ~Sample, colors = c("blue", "red", "grey"), 
        type = 'bar', barmode = 'group') %>%
  layout(title = "Comparison of League Avg vs Projected Lineup",
         xaxis = list(title = "Metrics"),
         yaxis = list(title = "Value"),
         legend = list(title = list(text = "Sample"))) 
  
  dis_df <- lineup_data %>%
  filter(pred_season > 2020)
league_dis <- dis_df %>%
  select(EfgPct,
         to_pct,
         off_reb_pct,
         ftr) %>%
  rename(`EFG%` = EfgPct,
         `FT Rate` = ftr,
         `OREB%` = off_reb_pct,
         `TO Rate` = to_pct)
euclidean_distances <- apply(league_dis, 1, function(row) {
  sqrt(sum((row - combined_preds[ ,-5])^2))
})
euc_res <- cbind(euclidean_distances, dis_df) %>%
  arrange(euclidean_distances) %>%
  head(5) %>%
  mutate(Year = pred_season - 1) %>%
  rename(Team = TeamAbbreviation,
         "Euclidean Distance" = euclidean_distances) %>%
  select(`Euclidean Distance`,
         Team,
         Year, 
         `Player 1`,
         `Player 2`,
         `Player 3`,
         `Player 4`,
         `Player 5`) 
  
  return(list(predictions = combined_preds[ , -5], chart = chart, similar_lineups = euc_res))
}

combined_function("Jaysom Tatum", "Jaylen Brown", "Kristaps Porzingis", "Derrick White", "Jrue Holiday")
```

shiny app - no optimization
```{r, error=FALSE, warning=FALSE, message=FALSE}
ui <- fluidPage(
  titlePanel("Basketball Lineup Projector"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("team", "Select Team:",
                  choices = unique(player_pool$team),
                  selected = NULL),
      uiOutput("player_select")
    ),
    
    mainPanel(
      tableOutput("lineup_preds"),  # Display predictions above the chart
      plotlyOutput("lineup_chart"),
      tableOutput("similar_lineups")
    )
  )
)

server <- function(input, output, session) {
  
  # Reactive expression to filter players based on selected team
  filtered_players <- reactive({
    req(input$team)
    player_pool %>%
      filter(team == input$team) %>%
      pull(namePlayer)
  })
  
  # Dynamic UI for player selection based on selected team
  output$player_select <- renderUI({
    req(filtered_players())
    checkboxGroupInput("players", "Select 5 Players:", 
                       choices = filtered_players(), 
                       selected = NULL,
                       inline = FALSE)
  })
  
  # Reactive expression to compute results based on selected players
  results <- reactive({
    req(input$players)
    if (length(input$players) != 5) {
      return(NULL)
    }
    combined_function(input$players[1], input$players[2], input$players[3], input$players[4], input$players[5])
  })
  
  # Render the predictions table
  output$lineup_preds <- renderTable({
    req(results())
    results()$predictions
  })
  
  # Render the plotly chart
  output$lineup_chart <- renderPlotly({
    req(results())
    results()$chart
  })
  
  # Render the similar lineups table
  output$similar_lineups <- renderTable({
    req(results())
    results()$similar_lineups
  })
}

# Run the Shiny app
shinyApp(ui = ui, server = server)

```

shiny app with optimization and exclusion box
```{r, error=FALSE, warning=FALSE, message=FALSE}
ui <- fluidPage(
  titlePanel("Basketball Lineup Projector"),
  
  # Add notes below the title panel using tags for HTML elements
  div(
    p("This app projects basketball lineups based on selected players and team data."),
    
    # Operation notes section
    p("Operation notes:"),
    tags$ul(
      tags$li("You can select 5 players to project their performance. Tha bar chart compares the lineup porjection with league and team averages."),
      tags$li("If fewer than 5 players are selected, the app will optimize the remaining players. The less `selected` players there are the longer the model takes to run"),
      tags$li("The optimization is based on one of the Four Factors: Effective Field Goal Percentage (EFG%), Free Throw Rate (FT Rate), Offensive Rebound Percentage (OREB%), or Turnover Rate (TO Rate).")
    ),
    
    # Model notes section
    p("Model notes:"),
    tags$ul(
      tags$li("The model is trained on lineup data from the last 10 seasons, limited to lineups that played at least 150 possessions."),
      tags$li("This is an ensemble model that combines a support vector machine model, gradient boosted model, and random forest model using a simple aggregation. This method leverages the principle of the wisdom of the crowd."),
      tags$li("Averages for the bar chart are from the last season.")
    ),
    
    style = "font-size:14px; color: #333;"
  ),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("team", "Select Team:",
                  choices = unique(player_pool$team),
                  selected = NULL),
      uiOutput("player_select"),            # Player selection
      uiOutput("exclude_player_select"),    # Exclude players box (now placed after player selection)
      selectInput("optimize_stat", "Optimize For Stat:",
                  choices = c("EFG%", "FT Rate", "OREB%", "TO Rate"),
                  selected = "EFG%"),
      actionButton("optimize_btn", "Run Model")  # Optimization button
    ),
    
    mainPanel(
      tableOutput("selected_players"),    # Display initially selected players
      tableOutput("optimized_players"),   # Display optimized players
      tableOutput("lineup_preds"),        # Display predictions
      plotlyOutput("lineup_chart"),       # Display chart
      uiOutput("similar_lineups_title"),  # Conditional title for similar lineups
      tableOutput("similar_lineups")      # Display similar lineups
    )
  )
)
server <- function(input, output, session) {
  
  # Reactive expression to filter players based on selected team
  filtered_players <- reactive({
    req(input$team)
    player_pool %>%
      filter(team == input$team) %>%
      pull(namePlayer)
  })
  
  # Dynamic UI for player selection based on selected team
  output$player_select <- renderUI({
    req(filtered_players())
    checkboxGroupInput("players", "Select Players:", 
                       choices = filtered_players(), 
                       selected = NULL,
                       inline = FALSE)
  })
  
  # Dynamic UI for exclude players based on selected team
  output$exclude_player_select <- renderUI({
    req(filtered_players())
    selectizeInput("exclude_players", "Exclude Players:", 
                   choices = filtered_players(), 
                   selected = NULL,
                   multiple = TRUE)
  })
  
  # Reactive expression to compute results based on selected players
  results <- eventReactive(input$optimize_btn, {
    req(input$players)
    
    # If exactly 5 players are selected, run the original function
    if (length(input$players) == 5) {
      combined_function(input$players[1], input$players[2], input$players[3], input$players[4], input$players[5])
      
    # If fewer than 5 players are selected, optimize to fill the lineup
    } else if (length(input$players) > 0) {
      num_missing <- 5 - length(input$players)
      selected_players <- input$players
      
      # Function to optimize for a given stat
      optimize_lineup <- function(stat) {
        all_players <- filtered_players()
        
        # Exclude the selected players and players to exclude from the available pool
        available_players <- setdiff(all_players, c(selected_players, input$exclude_players))
        
        player_combinations <- combn(available_players, num_missing, simplify = FALSE)
        
        best_score <- -Inf
        best_combination <- NULL
        
        for (combo in player_combinations) {
          lineup <- c(selected_players, combo)
          preds <- combined_function(lineup[1], lineup[2], lineup[3], lineup[4], lineup[5])
          score <- preds$predictions[[stat]]
          
          if (score > best_score) {
            best_score <- score
            best_combination <- combo
          }
        }
        
        return(best_combination)
      }
      
      # Optimize for the selected stat
      best_combination <- optimize_lineup(input$optimize_stat)
      final_players <- c(selected_players, best_combination)
      
      combined_results <- combined_function(final_players[1], final_players[2], final_players[3], final_players[4], final_players[5])
      
      return(list(
        selected = selected_players,
        optimized = best_combination,
        predictions = combined_results$predictions,
        chart = combined_results$chart,
        similar_lineups = combined_results$similar_lineups
      ))
      
    } else {
      return(NULL)
    }
  })
  
  # Render the initially selected players table
  output$selected_players <- renderTable({
    req(input$players)
    if (length(input$players) > 0) {
      data.frame("Selected Players" = input$players)
    } else {
      data.frame("Selected Players" = "None")
    }
  })
  
  # Render the optimized players table
  output$optimized_players <- renderTable({
    req(results())
    if (!is.null(results()$optimized)) {
      data.frame("Optimized Players" = results()$optimized)
    } else {
      data.frame("Optimized Players" = "None")
    }
  })
  
  # Render the predictions table
  output$lineup_preds <- renderTable({
    req(results())
    results()$predictions
  })
  
  # Render the plotly chart
  output$lineup_chart <- renderPlotly({
    req(results())
    results()$chart
  })
  
  # Conditionally render the similar lineups title
  output$similar_lineups_title <- renderUI({
    req(results())
    if (!is.null(results()$similar_lineups) && length(results()$similar_lineups) > 0) {
      h3("5 Similar Lineups from the Last 4 Seasons")
    } else {
      NULL
    }
  })
  
  # Render the similar lineups table
  output$similar_lineups <- renderTable({
  req(results())
  similar_lineups <- results()$similar_lineups
  
  # Format Year column
  if (!is.null(similar_lineups)) {
    similar_lineups$Year <- format(similar_lineups$Year, nsmall = 0)  # Ensure no decimal places
  }
  
  similar_lineups
})
}

# Run the Shiny app
shinyApp(ui = ui, server = server)
```

publishing app
```{r, error=FALSE, warning=FALSE, message=FALSE}
rsconnect::setAccountInfo(name='rbasketball',
			  token='C6402382F39CDAC8526C2D9323BABEB4',
			  secret='TwrPD0sQS/6Ni+CXrJuQab4LXPTf1VywskgfJ/wa')

deployApp('/Users/dylan/Desktop')
```

```{r, error=FALSE, warning=FALSE, message=FALSE}
getwd()

```

```{r, error=FALSE, warning=FALSE, message=FALSE}

```

```{r, error=FALSE, warning=FALSE, message=FALSE}

```

